---
title: "Lab 1"
subtitle: "Subsetting lists and base `R` `for()` loops"
execute: 
  echo: true
  eval: false
embed-resources: true
---

The purpose of this lab is to get you working with base `R` looping functions while also being careful about data types. 

You are welcome to work with a partner or in a small group of 3-4 people. 

If you include any content generated GenAI, you **must** cite it. See example in syllabus.

Please set `embed-resources: true` in the YAML header of quarto document so that your `html` output file can be displayed properly in a browser.


## Data structures

```{r}

mod <- lm(mpg ~ hp + cyl + drat, mtcars)

```

Run the above code

1. From the model results:
	a) Extract the qr tolerance (`tol`).
	b) Extract the "term labels" from the model.
		+ *Hint*: Use `str` to inspect the list. This one is tricky and involves pulling an attribute from an attribute.

```{r}
#| echo: false
#| eval: false
## 1.a
mod$qr$tol # qr tolerance

## 1.b
term_attributes <- attr(mod$model, "terms")
attr(term_attributes, "term.labels")
# - or - #
attr(mod$terms, "term.labels")
```

```{r}

set.seed(3000) # Set random number generation

l <- list(a = list(m = matrix(1:12, ncol = 3),
              	   v = 1:7),
          b = data.frame(student = 1:15,
                         score = rnorm(15, 100, 10)))
```

Run the above code

2. From `l`
	a) Use at least two different methods to extract `m`.
	b) Extract the third column of `m`. Maintain the dimensions (matrix structure).
	c) Extract the score for student 7 in the data frame.

```{r}
#| echo: false
#| eval: false

## 2.a
### Method 1
l$a$m

### Method 2
l[[1]][[1]]

l[["a"]][[1]]
l[["a"]][["m"]]

## 2.b
l$a$m[ ,3, drop = FALSE]

# as opposed to:
#l$a$m[ ,3]

## 2.c
### Method 1
l$b[7, ]

### Method 2
l[["b"]][ l[["b"]]$student == 7, ]
l[["b"]][7, ]

stu_ids <- l[["b"]]$student # these two lines are
l[["b"]][stu_ids == 7, ]    # equivalent to: l$b[7, ]
```

## Loops

3. Read in the `insurance_coverage.csv` file, which is in the data folder in the course repo. Split the data file by `"age_bucket"` and `"sex_name"`.

```{r}
#| echo: false
#| eval: false

file <- fs::dir_ls(here::here("data"), regexp = "insur")

insurance <- readr::read_csv(file)

by_age_sex <- split(
  insurance,
  list(insurance$age_bucket, insurance$sex_name)
)
```


4. Use a `for` loop to calculate the mean health care population (`hc_pop`) for each `"age_bucket"/"sex_name"` combination.

```{r}
#| echo: false
#| eval: false

mean_hc_pop1 <- rep(NA, length(by_age_sex))

# vector("list", length(by_age_sex)) # create an empty list instead of an atomic vector

for(i in seq_along(by_age_sex)) {
	mean_hc_pop1[i] <- mean(
	  by_age_sex[[i]]$hc_pop, 
	  na.rm = TRUE
  )
}
```

5. Replicate this calculation with `lapply`, `sapply`, and `vapply`

```{r}
#| echo: false
#| eval: false

## lapply
mean_hc_pop2 <- lapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE))

## sapply
mean_hc_pop3 <- sapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE))

## vapply
mean_hc_pop4 <- vapply(by_age_sex, function(df) mean(df$hc_pop, na.rm = TRUE),
                       FUN.VALUE = double(1))
```

6. Produce separate plots showing the change in `hc_pop` from 2014 to 2015 for each `"age_bucket"/"sex_name"` combination (with a single loop). Set the `hc_pop` axis so the limits equal the overall minimum and maximum values for `hc_pop`.

```{r}
#| echo: false
#| eval: false

theme_set(theme_minimal())

y_axis_range <- range(insurance$hc_pop)

plots <- lapply(by_age_sex, function(df) {
	ggplot(df, aes(year, hc_pop)) +
    geom_line(color = "gray60") +
    geom_point() +
    scale_x_continuous(breaks = c(2014, 2015)) +
    scale_y_continuous(limits = y_axis_range)
})

# extend this so it has a title on each
(titles <- names(by_age_sex))
(titles <- gsub("insurance_", "", titles))
(titles <- gsub("\\.", " ", titles))
(titles <- gsub("(\\d)to", "\\1 to ", titles))
(titles <- gsub("plus", " plus", titles))
(titles <- gsub("_", " ", titles))
(titles <- gsub("  ", " ", titles))

plots2 <- vector("list", length(by_age_sex))

for(i in seq_along(by_age_sex)) {
	plots2[[i]] <- 
  ggplot(by_age_sex[[i]], aes(year, hc_pop)) +
    geom_line(color = "gray60") +
    geom_point() +
    scale_x_continuous(breaks = c(2014, 2015)) +
    scale_y_continuous(limits = y_axis_range) +
    labs(title = titles[i])
}
```

7. Use a `for` loop to save the plots to a folder on your computer.

```{r}
#| echo: false
#| eval: false

dir.create(here::here("plots"))

filenames <- here::here(
  "plots",
  paste0(gsub("\\.", "-", names(by_age_sex)), ".png")
)

for(i in seq_along(plots)) {
	ggsave(filenames[i],
	       plots[[i]],
	       device = "png",
	       width = 6.5,
	       height = 8)
}
```

