---
title: "Lab 3"
subtitle: "Create and apply functions"
embed-resources: true
---

This lab focuses on writing basic functions. You are welcome to work with a partner or in a small group of 3-4 people. If you include any content generated GenAI, you **must** cite it. See example in syllabus.

Please set `embed-resources: true` in the YAML header of quarto document so that your `html` output file can be displayed properly in a browser. Please submit your `qmd` **and** html documents.

```{r}
#| include: false
library(tidyverse)
```

## Part A: Foundations

1. Briefly name and describe the three fundamental components of a function.

```{r}
#| label: ans-a1
#| include: FALSE

# 1. Formals - The arguments supplied to the function
# 2. Body - What the function actually does, including the use of the formals
# 3. Environment - Where the function "lives", which is important for scoping.
```

2. Describe three different ways functions can be stored/applied and an example of when you might want to use each version.

```{r}
#| label: ans-a2
#| include: FALSE

# anonymous functions are often used in loops, usually functionals like `purrr::map` or lapply.

# Functions are usually bound to a name and applied like we typically see functions, with the name preceding parentheses (e.g., `lm()`)

# Functions can also be bound to names within lists, which can be useful if you
# want to loop a set of functions through something, rather than looping a
# single function through a vector. In other words, the vector you are looping
# through is the list where each element of the list is a function
```

## Part B: Applied practice

The mean is defined as

$$
\frac{\sum_{i=1}^n{x_i}}{n}
$$

which basically says "sum the things and then divide by the total number of things".

1. Write a function to calculate the mean that removes missing data before conducting the calculation. You may not use `base::mean()` or any similar function that conducts a mean calculation. Include a `warning` in your function if missing data have been removed *that prints the total number of cases that were removed*.


```{r}
#| label: ans-b1
#| include: FALSE

my_mean <- function(x) {
	if(any(is.na(x))) {
		tot_na <- sum(is.na(x))
		warning(tot_na, " observations removed due to missingness")
		x <- na.omit(x)
		# x <- x[!is.na(x)]
	}
	sum(x) / length(x)
}

```

2. Test your function to make sure it:
    a) provides the expected results, and 
    b) gives identical output to `base::mean()` when `na.rm = TRUE`.
  
Test your function with `mtcars$mpg` and `airquality$Ozone`.

```{r}
#| label: ans-b2
#| include: FALSE

my_mean(mtcars$mpg)
mean(mtcars$mpg)

my_mean(airquality$Ozone)
mean(airquality$Ozone)
mean(airquality$Ozone, na.rm = TRUE)
```

3. Turn the following three lines of code into three different functions. Make sure to give them meaningful names. Test each of the functions to make sure they provide the expected output.

```{r}
#| label: practice
#| eval: false

# Count number of missing elements in a vector
sum(is.na(airquality$Ozone))

# Proportional representation of each level (unique element) in a vector
purrr::map_dbl(split(mtcars$cyl, mtcars$cyl), length) / length(mtcars$cyl)

# normalize or z-score a vector so the mean is zero and sd is one
(mtcars$mpg - mean(mtcars$mpg, na.rm = TRUE)) / sd(mtcars$mpg, na.rm = TRUE) 
```

```{r}
#| label: ans-b3
#| include: FALSE

count_missing <- function(x) {
	sum(is.na(x))
}
# test it
count_missing(c(rep(NA, 25), rnorm(75)))
count_missing(c(rep(NA, 42), rnorm(58)))

prop_categories <- function(cat_v) {
	purrr::map_dbl(split(cat_v, cat_v), length) / length(cat_v)
}
# test it
prop_categories(rep(letters[1:4], each = 25))
prop_categories(c(rep(letters[1], 50), rep(letters[4:5], each = 25)))

fix_missingness <- function(x) {
  if(any(is.na(x))) {
		tot_na <- sum(is.na(x))
		warning(paste(tot_na, "observations removed due to missingness"))
		x <- na.omit(x)
  }
  x
}

standardize <- function(x) {
	x <- fix_missingness(x)
	(x - mean(x)) / sd(x) 
}

# test it
test_standardize <- cbind(
  standardize(airquality$Ozone),
  na.omit(scale(airquality$Ozone))
)

identical(test_standardize[, 1], test_standardize[, 2])
```

4. Write a function that takes a data frame as its input and returns a data frame with only the numeric columns. Test that it works with the `forcats::gss_cat` dataset. 

```{r}
#| label: ans-b4
#| include: FALSE

extract_numeric_df <- function(df, as_tibble = TRUE) {
	num <- purrr::map_lgl(df, is.numeric)
	
	if(as_tibble) {
		return(tibble::as_tibble(df[, num]))
	}
	as.data.frame(df[, num])
}

extract_numeric_df(forcats::gss_cat)
extract_numeric_df(forcats::gss_cat, as_tibble = FALSE)
```

5. Write a function that uses the function you wrote in Question 4, and returns a data frame with the mean and standard deviation of each numeric column. If columns are omitted (because they are non-numeric), print a `message` stating the specific columns that were not included. Test that it works with the `palmerpenguins::penguins` and `forcats::gss_cat` datasets.

```{r}
#| label: ans-b51
#| include: false

df_summary <- function(df, as_tibble = TRUE) {
  df_num <- extract_numeric_df(df)
  means <- purrr::map_dbl(df_num, my_mean)
  
  sds <- purrr::map_dbl(df_num, sd, na.rm = TRUE)
  
  if(length(names(df)) != length(names(df_num))) {
    message("The following columns were not included in the summary: ", 
            glue::glue_collapse(setdiff(names(df), names(df_num)), sep = ", ", last = ", and ")) 
  }

  if (as_tibble) {
    tbl <- tibble::tibble(variable = names(df_num),
                          mean = means,
                          sd = sds)
  } else {
    tbl <- data.frame(variable = names(df_num),
                      mean = means,
                      sd = sds)
  }
  tbl
}
```


```{r}
#| label: ans-b52
#| include: FALSE

quiet_mymean <- purrr::quietly(my_mean) #capture side-effects (result/output AND warnings/messages)

df_summary <- function(df, as_tibble = TRUE) {
	df_num <- extract_numeric_df(df)
	means <- purrr::map(df_num, quiet_mymean)
	
	warns <- map(means, ~pluck(.x, "warnings")) |> 
    discard(~length(.x) == 0) |> 
    as_vector()
	warns <- gsub("^\\d+\\so(.+$)", "O\\1", warns)
	warns <- unique(warns[!is.na(warns)])
	
	means <- purrr::map_dbl(means, "result")
	  
	sds <- purrr::map_dbl(df_num, sd, na.rm = TRUE)
  
	if(length(names(df)) != length(names(df_num))) {
	  message("The following columns were not included in the summary: ", 
	          glue::glue_collapse(setdiff(names(df), names(df_num)), sep = ", ", last = ", and ")) 
	          #paste(setdiff(names(df), names(df_num)), collapse = ", ")) 
	}
	if(length(warns) > 0) {
	  warning(warns, call. = FALSE)
	}
	
	if (as_tibble) {
		tbl <- tibble::tibble(variable = names(df_num),
		                      mean = means,
		                      sd = sds)
	} else {
		tbl <- data.frame(variable = names(df_num),
		                  mean = means,
		                  sd = sds)
	}
	tbl
}

df_summary(palmerpenguins::penguins)
df_summary(forcats::gss_cat)
```

